---
title: "Angle Reference Table"
subtitle: "Angle between height and base for various rectangle proportions, And Your Own 'Quiver of Arrows'"
description: ""
authors: 
  - name: "Dan Swart, CPA (ret)"
date-format: long
# bibliography: manual-refs.bib
format:
  html:
    resources:
      - reference-backlinks.js
    include-after-body:    
      - text: |
          # <script type="text/javascript" src="reference-backlinks.js"></script>
    default: true         
    code-copy: true
    code-link: true        # This adds individual buttons
    code-fold: true        # Hide code by default, show on click
    code-summary: "Show the code"
    code-overflow: wrap
    code-block-bg: "#FAEBD7"
    code-block-border-left: "#31BAE9"
    embed-resources: true     # Set to 'false' for drafts. Override if you're NOT sharing output
    include-in-header:
      - text: 
          <link href="https://fonts.googleapis.com/css2?family=Lato&display=swap" rel="stylesheet">
          <link href="https://fonts.googleapis.com/css2?family=Fira+Mono&display=swap" rel="stylesheet">
          <link href="https://fonts.googleapis.com/css2?family=Montserrat&display=swap" rel="stylesheet">
          <link href="https://fonts.googleapis.com/css2?family=Cabin&display=swap" rel="stylesheet">
          <link href="https://fonts.googleapis.com/css2?family=Roboto+Slab&display=swap" rel="stylesheet">
          <link href="https://fonts.googleapis.com/css2?family=Schoolbell&display=swap" rel="stylesheet">
      - header.html
    css:
      - swart.css
      - tachyons.min.css
      - r-colors.css
    fontsize: 18pt
    lightbox: true
    page-layout: full
    fig-width: 13
    fig-height: 8
    fig-dpi: 300
    html-math-method: katex
    df-print: paged
    toc: false
    toc-float: true
    citeproc: true
    link-citations: true
    linestretch: 1.0
    
    
    
  typst:
    fig-width: 13
    fig-height: 8
    fig-dpi: 300
    margin:
      x: 1in
      y: 1in
    toc: false
    fontsize: 16pt
    mainfont: "Cabin"
  

    
  revealjs:
    slide-number: true
    transition: fade
    code-overflow: wrap
    center: true
    smaller: true
    scrollable: true
    chalkboard: true
    multiplex: false
    theme: solarized
    reference-location: margin
    logo: img/red-cross-640-435.png
    footer: "Footer text"
    code-block-height: 650px



  # docx:
  #   highlight-style: github
  #   fig_caption: true



editor: source

quarto:
  render:
    cache-refresh: true


# for .qmd filesd
execute:
  echo: false
  message: false
  warning: false
  eval: true
  fig-width: 13
  fig-height: 8


# for .rmd files
knitr:
  opts_chunk:
    echo: false
    error: false
    warning: false
    message: false
    cache: false


---


```{r}
#| label: setup
#| include: false


# install.packages(c("mapview", "survey", "srvyr", "arcgislayers"))

# census_api_key("95496766c51541ee6f402c1e1a8658581285b759", install = TRUE, overwrite = TRUE)


# # load libraries - NOT NEEDED


# Force dplyr's select to take precedence
select <- dplyr::select
filter <- dplyr::filter

# Options
options(scipen = 999)
options(qic.clshade = T) # NO LONGER NEEDED; CHARTS ALL PREPARED WITH GGPLOT2 ONLY
options(qic.linecol = 'black') # NO LONGER NEEDED; CHARTS ALL PREPARED WITH GGPLOT2 ONLY
options(qic.signalcol = "firebrick") # NO LONGER NEEDED; CHARTS ALL PREPARED WITH GGPLOT2 ONLY
options(qic.targetcol = "purple") # NO LONGER NEEDED; CHARTS ALL PREPARED WITH GGPLOT2 ONLY
options(DT.options = list(dom = 'pBlfrti')) # Add buttons, filtering, and top (t) pagination controls
options(shiny.maxRequestSize = 50 * 1024^2) # Set upload maximum to 50 MB
options(tigris_use_cache = TRUE)
options(device = "RStudioGD") 


# Flextable defaults:
flextable::set_flextable_defaults(
  font.size = 14, 
  font.family = "Cabin",
  font.color = "black",
  table.layout = "fixed",
  border.color = "darkgray",
  padding.top = 3, padding.bottom = 3,
  padding.left = 4, padding.right = 4,
  line_spacing = 1.3,
  digits = 2,
  decimal.mark = ".",
  big.mark = " ",
  na_str = "<na>",
  post_process_html = identity,
  post_process_docx = identity
)




# Set global theme for consistent plots
ggplot2::theme_set(
  ggplot2::theme_minimal(base_size = 20) +
    ggplot2::theme(
      plot.title.position = "plot",
      plot.title = ggtext::element_textbox_simple(
        family = "Cabin",
        face = "bold",
        color = "darkgreen",
        size = 16,
        fill = "yellow",
        lineheight = 0.9,
        padding = ggplot2::margin(5.5, 5.5, 0.0, 5.5),
        margin = ggplot2::margin(0, 0, 5.5, 0)
      ),
      plot.subtitle = ggtext::element_textbox_simple(
        family = "Cabin",
        color = "darkgreen",
        face = "bold",
        size = 14,
        fill = "yellow",
        lineheight = 0.9,
        padding = ggplot2::margin(5.5, 5.5, 5.5, 5.5),
        margin = ggplot2::margin(0.0, 0, 5.5, 0)
      ),
      plot.caption = ggtext::element_markdown(
        family = "Cabin",
        size = 14,
        hjust = 1,
        color = "darkblue",
        face = "italic",
        fill = "yellow",
        lineheight = 1.0
      ),
      axis.text.x = ggtext::element_markdown(
        family = "Cabin",
        face = "bold",
        color = "blue",
        size = 12,
        angle = 45,
        hjust = 1
      ),
        # ggplot2::element_blank(),
      axis.title.x = ggtext::element_markdown(
        family = "Cabin",
        face = "bold",
        color = "blue",
        size = 14),
        # ggplot2::element_blank(),
      axis.text.y = ggtext::element_markdown(
        family = "Cabin",
        face = "bold",
        color = "blue",
        size = 12,
        angle = 45,
        hjust = 1
      ),
        # ggplot2::element_blank(),
      axis.title.y = ggtext::element_markdown(
        family = "Cabin",
        face = "bold",
        color = "blue",
        size = 12),
        # ggplot2::element_blank(),
      strip.text = ggtext::element_markdown(
        family = "Cabin",
        color = "black",
        size = 12,
        face = "italic",
        margin = ggplot2::margin(2, 0, 0.5, 0, "lines")
      ),
      axis.text = ggtext::element_markdown(
        family = "Cabin",
        color = "black"),
      panel.background = ggplot2::element_rect(fill = "white", color = NA),
      plot.background = ggplot2::element_rect(fill = "white", color = NA),
      legend.position = "none",
      panel.spacing.x = grid::unit(1.5, "cm"),
      panel.spacing.y = grid::unit(1.5, "cm"),
      plot.margin = ggplot2::margin(20, 20, 20, 20, "pt")
    )
)



  
# Set seed for reproducibility
base::set.seed(123)

```

```{r}
#| label: angle-setup
#| message: false
sqrt_phi <- sqrt((1 + sqrt(5)) / 2)
phi      <- (1 + sqrt(5)) / 2
angles_df <- data.frame(
  base_name = c("√1", "√φ", "√2", "√2.25", "φ", "√3", "√4", "√5", "√6", "√7", "√9"),
  base_value = c(
    sqrt(1),
    sqrt_phi,
    sqrt(2),
    sqrt(2.25),
    phi,
    sqrt(3),
    sqrt(4),
    sqrt(5),
    sqrt(6),
    sqrt(7),
    sqrt(9)
  ),
  height = 1
)
angles_df$theta_rad <- atan2(angles_df$height, angles_df$base_value)
angles_df$theta_deg <- angles_df$theta_rad * (180 / pi)
display_df <- data.frame(
  Base = angles_df$base_name,
  `Base Length` = round(angles_df$base_value, 4),
  Height = 1,
  `Angle (degrees)` = round(angles_df$theta_deg, 2),
  check.names = FALSE
)
flextable::flextable(display_df) |>
  flextable::add_header_lines("Height = 1 for all rows") |>
  flextable::color(i = 1, part = "header", color = "blue") |>
  flextable::italic(i = 1, part = "header") |>
  flextable::align(i = 1, part = "header", align = "center") |>
  flextable::fontsize(i = 1, part = "header", size = 16) |>
  flextable::bg(i = 1, part = "header", bg = "white") |>
  flextable::bg(i = 2, part = "header", bg = "white") |>
  flextable::autofit()
```

\

***REVISED 2026-03-19***


Mount your printed 'Quiver of Arrows' under your tracing paper and draw the vertical and horizontal lines and extend each arrow to a length that will cover the largest image you will be analyzing.  You should be able to easily see the template beneath your tracing paper.

It is recommended you color code each arrow using the color coding of our fearless leader, Dan T.

You may need to trim the left and bottom edges of your tracing paper 'quiver' so it fits easily in the spine of a book.

It can be used to INSPECT existing work, or CONSTRUCT a framework for composing a new work.  It must be set in the lower LEFT corner of a work in landscape orientation, or the top LEFT corner for a work in portrait orientation.

The "√2.25" rectangle is also known as the "1.5 Rectangle".  The square root of 2.25 is 1.50.

\
\

::: {style="text-align: center;"}

**********
:::




{{< pagebreak >}}

```{=typst}
#set page(flipped: true)
```


```{r}
#| label: angle-graphic
#| fig-width: 13
#| fig-height: 8
#| out-width: "100%"

max_height <- 6
max_base <- 10
line_length <- sqrt(max_height^2 + max_base^2)
arrow_extension <- 2
lines_df <- data.frame(
  label = angles_df$base_name,
  theta_rad = angles_df$theta_rad
)
lines_df$xend <- line_length * cos(lines_df$theta_rad)
lines_df$yend <- line_length * sin(lines_df$theta_rad)
lines_df$xend <- pmin(lines_df$xend, max_base)
lines_df$yend <- ifelse(lines_df$xend == max_base, 
                        max_base * tan(lines_df$theta_rad), 
                        lines_df$yend)
lines_df$yend <- pmin(lines_df$yend, max_height)
lines_df$xend <- ifelse(lines_df$yend == max_height,
                        max_height / tan(lines_df$theta_rad),
                        lines_df$xend)

# Extend every arrow 2 inches beyond the reference rectangle edge
lines_df$xend <- lines_df$xend + arrow_extension * cos(lines_df$theta_rad)
lines_df$yend <- lines_df$yend + arrow_extension * sin(lines_df$theta_rad)

label_offset <- 0.6

# Default: push label outward along the angle from the new arrow tip
lines_df$label_x <- lines_df$xend + label_offset * cos(lines_df$theta_rad)
lines_df$label_y <- lines_df$yend + label_offset * sin(lines_df$theta_rad)

# Nudge a few crowded end labels 1 inch to the left
lines_df$label_x[lines_df$label %in% c("√φ", "√2", "√2.25", "φ")] <-
  lines_df$label_x[lines_df$label %in% c("√φ", "√2", "√2.25", "φ")] - 1

# Lower √3 through √9 labels by 0.50 inches
lines_df$label_y[lines_df$label %in% c("√3", "√4", "√5", "√6", "√7", "√9")] <-
  lines_df$label_y[lines_df$label %in% c("√3", "√4", "√5", "√6", "√7", "√9")] - 0.5

# Extend only the φ arrow another 1 inch, then move its label with it
lines_df$xend[lines_df$label == "φ"] <-
  lines_df$xend[lines_df$label == "φ"] + 1 * cos(lines_df$theta_rad[lines_df$label == "φ"])
lines_df$yend[lines_df$label == "φ"] <-
  lines_df$yend[lines_df$label == "φ"] + 1 * sin(lines_df$theta_rad[lines_df$label == "φ"])
lines_df$label_x[lines_df$label == "φ"] <-
  lines_df$xend[lines_df$label == "φ"] + label_offset * cos(lines_df$theta_rad[lines_df$label == "φ"]) - 1
lines_df$label_y[lines_df$label == "φ"] <-
  lines_df$yend[lines_df$label == "φ"] + label_offset * sin(lines_df$theta_rad[lines_df$label == "φ"])

# label_offset <- 0.6
# lines_df$label_x <- lines_df$xend + label_offset * cos(lines_df$theta_rad)
# lines_df$label_y <- lines_df$yend + label_offset * sin(lines_df$theta_rad)
ggplot2::ggplot() +
  ggplot2::geom_segment(
    data = lines_df,
    ggplot2::aes(x = 0, y = 0, xend = xend, yend = yend),
    arrow = ggplot2::arrow(length = ggplot2::unit(0.15, "inches"), type = "closed"),
    linewidth = 4.0
  ) +
  ggplot2::geom_text(
    data = lines_df,
    ggplot2::aes(x = label_x, y = label_y, label = label),
    hjust = 0, vjust = 0.5, size = 10, fontface = "bold"
  ) +
  ggplot2::geom_segment(
    ggplot2::aes(x = 0, y = 0, xend = 0, yend = max_height),
    linewidth = 4.0
  ) +
  ggplot2::geom_segment(
    ggplot2::aes(x = 0, y = 0, xend = max_base, yend = 0),
    linewidth = 4.0
  ) +
  ggplot2::coord_fixed(ratio = 1, xlim = c(-0.5, max(lines_df$label_x) + 0.5), ylim = c(-0.5, max(lines_df$label_y) + 0.5)) +
  ggplot2::theme_minimal() +
  ggplot2::labs(title = "Your 'Quiver of Arrows'") +
  ggplot2::theme(
    plot.title = ggplot2::element_text(size = 22, face = "bold"),
    axis.title = ggplot2::element_blank(),
    axis.text = ggplot2::element_blank(),
    axis.ticks = ggplot2::element_blank(),
    panel.grid = ggplot2::element_blank()
  )
```





```{=typst}
#set page(flipped: false)
```

