---
title: "Angle Reference Table"
subtitle: "Angle between height and base for various rectangle proportions, And Your Own Quiver of Arrows"
description: ""
authors: 
  - name: "Dan Swart, CPA (ret)"
date: today
date-format: long
# bibliography: manual-refs.bib
format:
  html:
    resources:
      - reference-backlinks.js
    include-after-body:    
      - text: |
          # <script type="text/javascript" src="reference-backlinks.js"></script>
    default: true         
    code-copy: true
    code-link: true        # This adds individual buttons
    code-fold: true        # Hide code by default, show on click
    code-summary: "Show the code"
    code-overflow: wrap
    code-block-bg: "#FAEBD7"
    code-block-border-left: "#31BAE9"
    embed-resources: true     # Set to 'false' for drafts. Override if you're NOT sharing output
    include-in-header:
      - text: 
          <link href="https://fonts.googleapis.com/css2?family=Lato&display=swap" rel="stylesheet">
          <link href="https://fonts.googleapis.com/css2?family=Fira+Mono&display=swap" rel="stylesheet">
          <link href="https://fonts.googleapis.com/css2?family=Montserrat&display=swap" rel="stylesheet">
          <link href="https://fonts.googleapis.com/css2?family=Cabin&display=swap" rel="stylesheet">
          <link href="https://fonts.googleapis.com/css2?family=Roboto+Slab&display=swap" rel="stylesheet">
          <link href="https://fonts.googleapis.com/css2?family=Schoolbell&display=swap" rel="stylesheet">
      - header.html
    css:
      - swart.css
      - tachyons.min.css
      - r-colors.css
    fontsize: 18pt
    lightbox: true
    page-layout: full
    fig-width: 13
    fig-height: 8
    fig-dpi: 300
    html-math-method: katex
    df-print: paged
    toc: false
    toc-float: true
    citeproc: true
    link-citations: true
    linestretch: 1.0
    
    
    
  typst:
    fig-width: 13
    fig-height: 8
    fig-dpi: 300
    margin:
      x: 1in
      y: 1in
    toc: false
    fontsize: 16pt
    mainfont: "Cabin"
  

    
  revealjs:
    slide-number: true
    transition: fade
    code-overflow: wrap
    center: true
    smaller: true
    scrollable: true
    chalkboard: true
    multiplex: false
    theme: solarized
    reference-location: margin
    logo: img/red-cross-640-435.png
    footer: "Footer text"
    code-block-height: 650px



  # docx:
  #   highlight-style: github
  #   fig_caption: true



editor: source

quarto:
  render:
    cache-refresh: true


# for .qmd filesd
execute:
  echo: false
  message: false
  warning: false
  eval: true
  fig-width: 13
  fig-height: 8


# for .rmd files
knitr:
  opts_chunk:
    echo: false
    error: false
    warning: false
    message: false
    cache: false


---


```{r}
#| label: setup
#| include: false


# install.packages(c("mapview", "survey", "srvyr", "arcgislayers"))

# census_api_key("95496766c51541ee6f402c1e1a8658581285b759", install = TRUE, overwrite = TRUE)


# # load libraries - NOT NEEDED


# Force dplyr's select to take precedence
select <- dplyr::select
filter <- dplyr::filter

# Options
options(scipen = 999)
options(qic.clshade = T) # NO LONGER NEEDED; CHARTS ALL PREPARED WITH GGPLOT2 ONLY
options(qic.linecol = 'black') # NO LONGER NEEDED; CHARTS ALL PREPARED WITH GGPLOT2 ONLY
options(qic.signalcol = "firebrick") # NO LONGER NEEDED; CHARTS ALL PREPARED WITH GGPLOT2 ONLY
options(qic.targetcol = "purple") # NO LONGER NEEDED; CHARTS ALL PREPARED WITH GGPLOT2 ONLY
options(DT.options = list(dom = 'pBlfrti')) # Add buttons, filtering, and top (t) pagination controls
options(shiny.maxRequestSize = 50 * 1024^2) # Set upload maximum to 50 MB
options(tigris_use_cache = TRUE)
options(device = "RStudioGD") 


# Flextable defaults:
flextable::set_flextable_defaults(
  font.size = 14, 
  font.family = "Cabin",
  font.color = "black",
  table.layout = "fixed",
  border.color = "darkgray",
  padding.top = 3, padding.bottom = 3,
  padding.left = 4, padding.right = 4,
  line_spacing = 1.3,
  digits = 2,
  decimal.mark = ".",
  big.mark = " ",
  na_str = "<na>",
  post_process_html = identity,
  post_process_docx = identity
)



# Flextable built-in themes:
  # flextable::theme_alafoli()	|>  # BLAH
  # flextable::theme_apa()  # THIS IS NICE
  # flextable::theme_booktabs() |>  # NICE, MORE COMPACT
  # flextable::theme_box() |>   # OK, INCLUDES CELL BORDERS
  # flextable::theme_tron() |>  # 'DARK MODE' BLUE TEXT
  # flextable::theme_tron_legacy() |>   # 'DARK MODE' YELLOW TEXT
  # flextable::theme_vader() |>    # 'DARK MODE' WHITE TEXT
  # flextable::theme_vanilla() |>   # NOT SPECIAL
  # flextable::theme_zebra()	|>
  #


# Set global theme for consistent plots
ggplot2::theme_set(
  ggplot2::theme_minimal(base_size = 20) +
    ggplot2::theme(
      plot.title.position = "plot",
      plot.title = ggtext::element_textbox_simple(
        family = "Cabin",
        face = "bold",
        color = "darkgreen",
        size = 16,
        fill = "yellow",
        lineheight = 0.9,
        padding = ggplot2::margin(5.5, 5.5, 0.0, 5.5),
        margin = ggplot2::margin(0, 0, 5.5, 0)
      ),
      plot.subtitle = ggtext::element_textbox_simple(
        family = "Cabin",
        color = "darkgreen",
        face = "bold",
        size = 14,
        fill = "yellow",
        lineheight = 0.9,
        padding = ggplot2::margin(5.5, 5.5, 5.5, 5.5),
        margin = ggplot2::margin(0.0, 0, 5.5, 0)
      ),
      plot.caption = ggtext::element_markdown(
        family = "Cabin",
        size = 14,
        hjust = 1,
        color = "darkblue",
        face = "italic",
        fill = "yellow",
        lineheight = 1.0
      ),
      axis.text.x = ggtext::element_markdown(
        family = "Cabin",
        face = "bold",
        color = "blue",
        size = 12,
        angle = 45,
        hjust = 1
      ),
        # ggplot2::element_blank(),
      axis.title.x = ggtext::element_markdown(
        family = "Cabin",
        face = "bold",
        color = "blue",
        size = 14),
        # ggplot2::element_blank(),
      axis.text.y = ggtext::element_markdown(
        family = "Cabin",
        face = "bold",
        color = "blue",
        size = 12,
        angle = 45,
        hjust = 1
      ),
        # ggplot2::element_blank(),
      axis.title.y = ggtext::element_markdown(
        family = "Cabin",
        face = "bold",
        color = "blue",
        size = 12),
        # ggplot2::element_blank(),
      strip.text = ggtext::element_markdown(
        family = "Cabin",
        color = "black",
        size = 12,
        face = "italic",
        margin = ggplot2::margin(2, 0, 0.5, 0, "lines")
      ),
      axis.text = ggtext::element_markdown(
        family = "Cabin",
        color = "black"),
      panel.background = ggplot2::element_rect(fill = "white", color = NA),
      plot.background = ggplot2::element_rect(fill = "white", color = NA),
      legend.position = "none",
      panel.spacing.x = grid::unit(1.5, "cm"),
      panel.spacing.y = grid::unit(1.5, "cm"),
      plot.margin = ggplot2::margin(20, 20, 20, 20, "pt")
    )
)



  
# Set seed for reproducibility
base::set.seed(123)

```

```{r}
#| label: angle-setup
#| message: false
sqrt_phi <- sqrt((1 + sqrt(5)) / 2)
phi      <- (1 + sqrt(5)) / 2
angles_df <- data.frame(
  base_name = c("√1", "√φ", "√2", "φ", "√3", "√4", "√5", "√6", "√7"),
  base_value = c(sqrt(1), sqrt_phi, sqrt(2), phi, sqrt(3), sqrt(4), sqrt(5), sqrt(6), sqrt(7)),
  height = 1
)
angles_df$theta_rad <- atan2(angles_df$height, angles_df$base_value)
angles_df$theta_deg <- angles_df$theta_rad * (180 / pi)
display_df <- data.frame(
  Base = angles_df$base_name,
  `Base Length` = round(angles_df$base_value, 4),
  Height = 1,
  `Angle (degrees)` = round(angles_df$theta_deg, 2),
  check.names = FALSE
)
flextable::flextable(display_df) |>
  flextable::add_header_lines("Height = 1 for all rows") |>
  flextable::color(i = 1, part = "header", color = "blue") |>
  flextable::italic(i = 1, part = "header") |>
  flextable::align(i = 1, part = "header", align = "center") |>
  flextable::fontsize(i = 1, part = "header", size = 16) |>
  flextable::bg(i = 1, part = "header", bg = "white") |>
  flextable::bg(i = 2, part = "header", bg = "white") |>
  flextable::autofit()
```

\
Your 'Quiver of Arrows' below is made to analyze any rectangle from left to right.  It can be used to INSPECT existing work, or CONSTRUCT a framework for composing a new work.  This means it will only work when you place it in the lower left corner, or the top right corner with the long side as the base.  If you insist on using it on the bottom right or top left the root2 arrow will point exactly to the midpoint of the opposite side.  Try it, you'll see.  It's an extra bonus that you get for free!

\
\

::: {style="text-align: center;"}

**********
:::

I know it works on landscape orientation, but not sure yet if it works in portrait orientation.  I haven't gotten that far in the lecture.

I recommend placing it at the bottom left corner.  Short side = the height, longer side is the base.

Slip it under the tracing paper to verify the construction of your work, look through the traccing paper to ANALIZE an existing image, or use it to draw the correct diagonal for the appropriate rectangle.


{{< pagebreak >}}

```{=typst}
#set page(flipped: true)
```


```{r}
#| label: angle-graphic
#| fig-width: 13
#| fig-height: 8
#| out-width: "100%"

max_height <- 6
max_base <- 10
line_length <- sqrt(max_height^2 + max_base^2)
lines_df <- data.frame(
  label = angles_df$base_name,
  theta_rad = angles_df$theta_rad
)
lines_df$xend <- line_length * cos(lines_df$theta_rad)
lines_df$yend <- line_length * sin(lines_df$theta_rad)
lines_df$xend <- pmin(lines_df$xend, max_base)
lines_df$yend <- ifelse(lines_df$xend == max_base, 
                        max_base * tan(lines_df$theta_rad), 
                        lines_df$yend)
lines_df$yend <- pmin(lines_df$yend, max_height)
lines_df$xend <- ifelse(lines_df$yend == max_height,
                        max_height / tan(lines_df$theta_rad),
                        lines_df$xend)
label_offset <- 0.6

# Detect which lines are clamped to bounding box edges
hit_top   <- abs(lines_df$yend - max_height) < 0.01
hit_right <- abs(lines_df$xend - max_base)   < 0.01

# Default: push label outward along the angle
lines_df$label_x <- lines_df$xend + label_offset * cos(lines_df$theta_rad)
lines_df$label_y <- lines_df$yend + label_offset * sin(lines_df$theta_rad)

# Lines hitting the ceiling: push label straight UP so closely-spaced
# angles (like sqrt(phi) and sqrt(2)) don't stack on the same y-position
lines_df$label_x[hit_top] <- lines_df$xend[hit_top]
lines_df$label_y[hit_top] <- lines_df$yend[hit_top] + label_offset

# Lines hitting the right wall: push label straight RIGHT
lines_df$label_x[hit_right] <- lines_df$xend[hit_right] + label_offset
lines_df$label_y[hit_right] <- lines_df$yend[hit_right]


# label_offset <- 0.6
# lines_df$label_x <- lines_df$xend + label_offset * cos(lines_df$theta_rad)
# lines_df$label_y <- lines_df$yend + label_offset * sin(lines_df$theta_rad)
ggplot2::ggplot() +
  ggplot2::geom_segment(
    data = lines_df,
    ggplot2::aes(x = 0, y = 0, xend = xend, yend = yend),
    arrow = ggplot2::arrow(length = ggplot2::unit(0.15, "inches"), type = "closed"),
    linewidth = 4.5
  ) +
  ggplot2::geom_text(
    data = lines_df,
    ggplot2::aes(x = label_x, y = label_y, label = label),
    hjust = 0, vjust = 0.5, size = 10, fontface = "bold"
  ) +
  ggplot2::geom_segment(
    ggplot2::aes(x = 0, y = 0, xend = 0, yend = max_height),
    linewidth = 4.5
  ) +
  ggplot2::geom_segment(
    ggplot2::aes(x = 0, y = 0, xend = max_base, yend = 0),
    linewidth = 4.5
  ) +
  ggplot2::coord_fixed(ratio = 1, xlim = c(-0.5, max_base + 1), ylim = c(-0.5, max_height + 0.5)) +
  ggplot2::theme_minimal() +
  ggplot2::labs(title = "Your 'Quiver of Arrows'") +
  ggplot2::theme(
    plot.title = ggplot2::element_text(size = 22, face = "bold"),
    axis.title = ggplot2::element_blank(),
    axis.text = ggplot2::element_blank(),
    axis.ticks = ggplot2::element_blank(),
    panel.grid = ggplot2::element_blank()
  )
```





```{=typst}
#set page(flipped: false)
```

