---
title: "√2 Rectangle Arrow Tool: Why It Matches in Some Corners (and Why Midpoints Appear)"
author: "Dan Swart"
date: "2026-02-15"
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    code-tools: true
  pdf:
    toc: true
execute:
  echo: true
  warning: false
  message: false
---

## Setup

```{r}
#| label: setup
#| include: true

# Core
library(ggplot2)
library(dplyr)
library(tidyr)

# Tables
library(flextable)

# Optional niceties (safe if installed)
# library(scales)
```

## Problem statement (your construction)

You drew a rectangle with:

- Height \(h = 36\) cm  
- Base \(w = 36\sqrt{2} \approx 50.9\) cm  

You then created a “tool” (a fan of arrows) from a corner. Each arrow is drawn at an angle computed from the base, using **method B**:

\[
\theta_n = \arctan\!\left(\frac{1}{\sqrt{n}}\right)
\]

You used \(n \in \{\varphi, 2, 3, 4, 5, 6, 7\}\) (with \(\varphi\) the golden ratio) and expected the \(\sqrt{2}\) arrow to align with the drawn diagonal everywhere.

It **does** match perfectly in some placements, but **not for every corner**. Even more interesting: the “\(\sqrt2\) arrow” placed at certain corners points directly to the **midpoint** of an opposite side.

The geometry is doing exactly what it should—this is an **orientation / quadrant** issue plus a **complement-angle** phenomenon.

---

## Key facts

### The √2 rectangle diagonal angle (method B)

For a \(\sqrt2\) rectangle, \(w = h\sqrt2\). The diagonal from bottom-left has slope:

\[
\text{slope} = \frac{h}{w} = \frac{h}{h\sqrt2} = \frac{1}{\sqrt2}
\]

So the diagonal angle above the base (when the base points to the right) is:

\[
\theta = \arctan\!\left(\frac{1}{\sqrt2}\right)\approx 35.264^\circ
\]

This is exactly what you computed with **B**.

### Complement angle (the one that causes midpoints)

The complement is:

\[
90^\circ - \theta
= 90^\circ - \arctan\!\left(\frac{1}{\sqrt2}\right)
= \arctan(\sqrt2)
\approx 54.736^\circ
\]

That complementary direction has slope \(\sqrt2\) (steeper). In a \(\sqrt2\) rectangle, that ray has a special “midpoint” intersection property.

---

## Diagram 1 — The √2 rectangle and its true diagonal

```{r}
#| label: fig-rectangle-diagonal
#| fig-cap: "√2 rectangle (h = 36, w = 36√2) and the diagonal angle θ = arctan(1/√2) ≈ 35.264° measured from the base."
#| fig-width: 7
#| fig-height: 5

h <- 36
w <- h * sqrt(2)

theta <- atan(1/sqrt(2))          # radians
theta_deg <- theta * 180/pi

# Data for segments
df_seg <- tibble::tibble(
  x = c(0, 0),
  y = c(0, 0),
  xend = c(w, w),
  yend = c(0, h),
  which = c("base", "diagonal")
)

ggplot2::ggplot() +
  ggplot2::geom_rect(
    ggplot2::aes(xmin = 0, xmax = w, ymin = 0, ymax = h),
    fill = NA,
    linewidth = 0.7
  ) +
  ggplot2::geom_segment(
    data = df_seg,
    ggplot2::aes(x = x, y = y, xend = xend, yend = yend, linetype = which),
    linewidth = 0.9
  ) +
  ggplot2::scale_linetype_manual(values = c(base = "solid", diagonal = "solid")) +
  ggplot2::annotate("text", x = w/2, y = -1.5, label = "base (w = h√2)", vjust = 1) +
  ggplot2::annotate("text", x = w + 1.5, y = h/2, label = "height (h)", hjust = 0) +
  ggplot2::annotate(
    "text",
    x = w*0.35, y = h*0.16,
    label = paste0("θ = arctan(1/√2)\n≈ ", round(theta_deg, 3), "°"),
    hjust = 0
  ) +
  ggplot2::coord_equal(expand = TRUE) +
  ggplot2::theme_minimal(base_size = 12) +
  ggplot2::labs(x = NULL, y = NULL)
```

### What you’re seeing

From the **bottom-left corner**, measuring angle upward from the bottom edge:

- The diagonal makes about **35.264°** with the base.
- That matches your computed angle (method B).
- So your arrow aligns perfectly.

This is the “it works perfectly here” case.

---

## Why it fails at other corners (the quadrant/orientation issue)

The diagonal is one geometric direction, but the **numeric angle you must use depends on how the local base edge is oriented**.

If your tool assumes:

- the base runs left → right  
- and angles are measured counter-clockwise from that base

…then moving the tool to a corner where the edge direction is reversed (right → left), or where you’re effectively measuring from a vertical edge, requires angle transforms.

---

## Table — Correct diagonal direction at each corner

Let \(\theta \approx 35.264^\circ\).

```{r}
#| label: tbl-corner-angles
#| tbl-cap: "Same diagonal direction, but the numeric angle depends on the corner (quadrant)."

corner_tbl <- tibble::tibble(
  corner = c("Bottom-Left", "Bottom-Right", "Top-Right", "Top-Left"),
  correct_angle = c("θ",
                    "180° − θ",
                    "180° + θ",
                    "360° − θ"),
  numeric_degrees = c(theta_deg,
                      180 - theta_deg,
                      180 + theta_deg,
                      360 - theta_deg)
) |>
  dplyr::mutate(numeric_degrees = round(numeric_degrees, 3))

flextable::flextable(corner_tbl) |>
  flextable::colformat_num(j = "numeric_degrees", digits = 3) |>
  flextable::autofit()
```

---

## Diagram 2 — The diagonal direction from all corners

This plot draws the diagonal direction as a ray from each corner.

```{r}
#| label: fig-corners-diagonals
#| fig-cap: "Diagonal direction rays from all four corners. The direction is the same geometry, but the angle label changes by quadrant."
#| fig-width: 7
#| fig-height: 5

corners <- tibble::tibble(
  corner = c("BL","BR","TR","TL"),
  x0 = c(0, w, w, 0),
  y0 = c(0, 0, h, h),
  x1 = c(w, 0, 0, w),
  y1 = c(h, h, 0, 0)
)

ggplot2::ggplot() +
  ggplot2::geom_rect(
    ggplot2::aes(xmin = 0, xmax = w, ymin = 0, ymax = h),
    fill = NA,
    linewidth = 0.7
  ) +
  ggplot2::geom_segment(
    data = corners,
    ggplot2::aes(x = x0, y = y0, xend = x1, yend = y1),
    linewidth = 0.9
  ) +
  ggplot2::geom_point(
    data = corners,
    ggplot2::aes(x = x0, y = y0),
    size = 2.3
  ) +
  ggplot2::annotate("text", x = 0, y = 0, label = "BL", hjust = -0.2, vjust = -0.6) +
  ggplot2::annotate("text", x = w, y = 0, label = "BR", hjust = 1.2, vjust = -0.6) +
  ggplot2::annotate("text", x = w, y = h, label = "TR", hjust = 1.2, vjust = 1.3) +
  ggplot2::annotate("text", x = 0, y = h, label = "TL", hjust = -0.2, vjust = 1.3) +
  ggplot2::coord_equal(expand = TRUE) +
  ggplot2::theme_minimal(base_size = 12) +
  ggplot2::labs(x = NULL, y = NULL)
```

### The core insight

Your geometry is correct.

The mismatch is entirely about this:

> **Angle direction depends on edge orientation.**

The diagonal is a fixed line. But the **angle number** that describes it depends on which axis and orientation you’re measuring from.

---

## The midpoint phenomenon (the “beautiful surprise”)

You noticed something that feels strange at first:

> The √2 arrow, when placed at two of the other corners, points directly to the midpoint of the opposite side.

That midpoint behavior happens when the direction you actually draw corresponds to the **complement angle**:

\[
90^\circ-\theta = \arctan(\sqrt2)\approx 54.736^\circ
\]

That angle has slope \(\sqrt2\). In a \(\sqrt2\) rectangle, a ray with slope \(\sqrt2\) from a corner hits the opposite side exactly at its midpoint.

### Algebra (coordinate proof)

Let the rectangle have height \(h\) and width \(w=h\sqrt2\).

From the bottom-left corner, draw the line:

\[
y = \sqrt2\,x
\]

It hits the top edge when \(y=h\), so:

\[
h = \sqrt2\,x \quad \Rightarrow \quad x=\frac{h}{\sqrt2}
\]

But half the width is:

\[
\frac{w}{2} = \frac{h\sqrt2}{2} = \frac{h}{\sqrt2}
\]

So \(x=w/2\). That is **exactly the midpoint** of the top edge.

This is not a coincidence; it’s a built-in symmetry of \(\sqrt2\).

---

## Diagram 3 — The midpoint ray (slope √2) vs the diagonal (slope 1/√2)

```{r}
#| label: fig-midpoint-vs-diagonal
#| fig-cap: "In a √2 rectangle, the slope √2 ray hits the midpoint of the top edge; the true diagonal uses slope 1/√2."
#| fig-width: 7
#| fig-height: 5

# Intersection of y = sqrt(2) x with y = h
x_mid <- h / sqrt(2)

df_lines <- tibble::tibble(
  which = c("diagonal (slope 1/√2)", "midpoint ray (slope √2)"),
  x = c(0, 0),
  y = c(0, 0),
  xend = c(w, x_mid),
  yend = c(h, h)
)

df_mid <- tibble::tibble(
  x = x_mid,
  y = h,
  label = "midpoint of top edge"
)

ggplot2::ggplot() +
  ggplot2::geom_rect(
    ggplot2::aes(xmin = 0, xmax = w, ymin = 0, ymax = h),
    fill = NA,
    linewidth = 0.7
  ) +
  ggplot2::geom_segment(
    data = df_lines,
    ggplot2::aes(x = x, y = y, xend = xend, yend = yend, linetype = which),
    linewidth = 0.9
  ) +
  ggplot2::geom_point(
    data = df_mid,
    ggplot2::aes(x = x, y = y),
    size = 2.5
  ) +
  ggplot2::annotate("text", x = x_mid, y = h + 1.2, label = "midpoint", vjust = 0) +
  ggplot2::annotate("text", x = w*0.58, y = h*0.55,
                    label = "diagonal: tan(θ)=1/√2", hjust = 0) +
  ggplot2::annotate("text", x = w*0.25, y = h*0.80,
                    label = "complement: tan(90°−θ)=√2", hjust = 0) +
  ggplot2::coord_equal(expand = TRUE) +
  ggplot2::theme_minimal(base_size = 12) +
  ggplot2::labs(x = NULL, y = NULL) +
  ggplot2::scale_linetype_manual(values = c("diagonal (slope 1/√2)" = "solid",
                                            "midpoint ray (slope √2)" = "dashed"))
```

---

## Practical “place the tool at any corner” rule

Let your method-B angle be:

\[
\theta_n = \arctan\!\left(\frac{1}{\sqrt{n}}\right)
\]

When you place the tool at a new corner, you must account for the local base direction:

- If the base edge points **left → right**, you can use \(\theta_n\) directly (counter-clockwise from the base).
- If the base edge points **right → left**, use \(180^\circ - \theta_n\).
- If you’re measuring from a vertical edge instead of the base, use the complement \(90^\circ - \theta_n\) (or convert accordingly).

A lot of “it works here but not there” comes from needing a **mirror flip** (reflection) rather than a pure rotation.

---

## Extending the idea to √3, √5, √φ, etc.

Your method B angle for any \(n>0\):

\[
\theta_n = \arctan\!\left(\frac{1}{\sqrt{n}}\right)
\quad\Rightarrow\quad
\tan(\theta_n)=\frac{1}{\sqrt{n}}
\]

The complementary direction is:

\[
90^\circ-\theta_n = \arctan(\sqrt{n})
\quad\Rightarrow\quad
\tan(90^\circ-\theta_n)=\sqrt{n}
\]

So every time you see an “unexpected” but *consistent* targeting behavior (like midpoints), it’s a strong hint that you’re drawing a **complement** (slope \(\sqrt{n}\)) rather than the intended method-B direction (slope \(1/\sqrt{n}\)).

### Quick reference table (method B)

```{r}
#| label: tbl-general-angles
#| tbl-cap: "Method B angles θn = arctan(1/√n) and their complements."

phi <- (1 + sqrt(5))/2

n_vals <- tibble::tibble(
  label = c("φ", "2", "3", "4", "5", "6", "7"),
  n = c(phi, 2, 3, 4, 5, 6, 7)
) |>
  dplyr::mutate(
    theta_rad = atan(1/sqrt(n)),
    theta_deg = theta_rad * 180/pi,
    comp_deg  = 90 - theta_deg
  ) |>
  dplyr::mutate(
    theta_deg = round(theta_deg, 3),
    comp_deg  = round(comp_deg, 3)
  ) |>
  dplyr::select(label, n, theta_deg, comp_deg)

flextable::flextable(n_vals) |>
  flextable::set_header_labels(
    label = "n (label)",
    n = "n (value)",
    theta_deg = "θn = arctan(1/√n) (deg)",
    comp_deg = "90° − θn (deg)"
  ) |>
  flextable::autofit()
```

---

## One more diagnostic diagram: the “fan” from a corner

This last plot shows your idea: multiple method-B rays from the bottom-left corner.

```{r}
#| label: fig-fan
#| fig-cap: "A method-B fan of rays from the bottom-left corner for n ∈ {φ,2,3,4,5,6,7}. The √2 diagonal is included for reference."
#| fig-width: 7
#| fig-height: 5

phi <- (1 + sqrt(5))/2
fan <- tibble::tibble(
  label = c("φ", "2", "3", "4", "5", "6", "7"),
  n = c(phi, 2, 3, 4, 5, 6, 7)
) |>
  dplyr::mutate(
    theta = atan(1/sqrt(n)),      # radians
    # pick a common ray length to show direction
    L = w * 0.95,
    xend = L * cos(theta),
    yend = L * sin(theta)
  )

diag <- tibble::tibble(
  label = "√2 diagonal",
  x = 0, y = 0,
  xend = w, yend = h
)

ggplot2::ggplot() +
  ggplot2::geom_rect(
    ggplot2::aes(xmin = 0, xmax = w, ymin = 0, ymax = h),
    fill = NA,
    linewidth = 0.7
  ) +
  ggplot2::geom_segment(
    data = fan,
    ggplot2::aes(x = 0, y = 0, xend = xend, yend = yend),
    linewidth = 0.8,
    alpha = 0.85
  ) +
  ggplot2::geom_segment(
    data = diag,
    ggplot2::aes(x = x, y = y, xend = xend, yend = yend),
    linewidth = 1.0
  ) +
  ggplot2::geom_text(
    data = fan,
    ggplot2::aes(x = xend, y = yend, label = label),
    hjust = -0.1,
    vjust = 0.5,
    size = 3.5
  ) +
  ggplot2::coord_equal(expand = TRUE) +
  ggplot2::theme_minimal(base_size = 12) +
  ggplot2::labs(x = NULL, y = NULL)
```

---

## Summary (the narrative, in one place)

- With \(h=36\) and \(w=36\sqrt2\approx 50.9\), your rectangle is essentially a true \(\sqrt2\) rectangle.
- Using method **B**, the diagonal angle from the base is  
  \(\theta=\arctan(1/\sqrt2)\approx 35.264^\circ\).  
  That’s why it matches perfectly at the bottom-left corner when the base runs left → right.
- When you move the tool to other corners, the *geometry* is unchanged, but the **numeric angle label changes by quadrant**:
  \(\theta\), \(180^\circ-\theta\), \(180^\circ+\theta\), \(360^\circ-\theta\).
- The “midpoint” behavior is the complementary angle phenomenon:  
  \(90^\circ-\theta=\arctan(\sqrt2)\approx 54.736^\circ\).  
  In a \(\sqrt2\) rectangle, a slope-\(\sqrt2\) ray from a corner hits the opposite side at its midpoint exactly.
- If your tool is rotated but not mirrored (or if you measure from a different edge), you can accidentally draw the complement direction—producing consistent-but-surprising results like midpoint hits.
